\section{Oracle demonstrators}
\label{sec:oracles}

Two scripted oracles produce demonstrations. Both consume the same task
scripts (time-parameterised end-effector tracks and gripper schedules)
and both are gated by the success criterion of Section~\ref{sec:tasks};
they differ only in how end-effector targets become joint commands. The
\emph{teleoperation-pipeline} oracle is the default; the \emph{direct}
oracle is the deterministic fallback. This section explains why the
pipeline-authentic route matters, derives its inverse retargeting map,
describes the closed-loop corrections that stand in for the operator's
eyes, and reports the reliability both oracles reach.

\paragraph{Why demonstrate through the pipeline.}
The purpose of the simulated campaign is to de-risk a real one, and the
real one records \emph{teleoperated} data. The teleoperation stack
shapes that data everywhere: the adaptive input filter adds
state-dependent lag, the clutch re-anchors the mapping at every grip,
the retargeting scales and rotates operator motion, and the solver runs
asynchronously at its own rate, so commands interleave with observations
exactly as its thread timing dictates. A policy trained on
demonstrations that bypass all of this meets those dynamics for the
first time on the real rig. Our oracle therefore \emph{is} a scripted
operator: it implements the same reader interface as the headset,
holds both grips, moves virtual hands, and squeezes triggers, and the
production stack --- filter, clutch calibration, operator-frame
retargeting, asynchronous solver, rate limiting --- runs unmodified
between the oracle and the simulated arms.

\paragraph{Inverting the retargeting.}
At grip engagement the stack calibrates a map from hand space to robot
space: it stores the hand rest position $H_0$ and the arm's current
end-effector position $E_0$, and thereafter commands
\begin{equation}
  p(t) \;=\; E_0 + s\, R_{\mathrm{op}}\,\bigl(h(t) - H_0\bigr),
  \label{eq:retarget}
\end{equation}
where $h(t)$ is the filtered hand position, $s$ the translation scale
and $R_{\mathrm{op}}$ the operator control frame, a rotation the stack
computes deterministically from the rest pose. The oracle wants the
pipeline to output a scripted target $p^\ast(t)$, so it places its
virtual hand at the closed-form inverse of Equation~\ref{eq:retarget}:
\begin{equation}
  h(t) \;=\; H_0 + \tfrac{1}{s}\, R_{\mathrm{op}}^{\!\top}
  \bigl(p^\ast(t) - E_0\bigr).
\end{equation}
Hands keep the identity orientation (handles pointing down, the
calibration posture), and the trigger value is one minus the scripted
gripper fraction. The demonstration is then whatever the pipeline makes
of these inputs --- including its lag and filtering --- and the recorded
action is the joint command the solver actually published, matching the
real recorder's semantics. Collection consequently runs in real time and
is not perfectly repeatable; scenarios remain seeded.

\paragraph{Closed-loop corrections.}
The differential solver tracks with a steady-state positional error of
roughly \SI{8}{\milli\metre} that varies with the arm's azimuth. A
\SI{10}{\centi\metre} object shrugs this off; a
\SI{2.2}{\centi\metre} cube does not --- the jaws close beside it. A
human operator absorbs exactly this error visually, so the oracle gets
eyes: a \emph{grasp servo} nudges the approach target to centre the
pinch on the measured cube position while the jaws are still open, and
freezes the moment they begin to close; a \emph{hold servo} steers the
gripped cube onto the scripted track, and freezes once the cube is set
back down, because continuing to correct through the pipeline's lag
drags an already-placed object off its target. Both corrections use low
gains with hard magnitude clips: the pipeline lags the target by several
ticks, and a high-gain correction integrates that lag into wind-up that
throws the pinch off entirely (Section~\ref{sec:difficulties}).
\unjustified{} the gains, clips and dwell durations were tuned by
inspecting rendered episodes; they are stable but not derived.

\paragraph{The direct oracle.}
The fallback path feeds the scripted targets straight into the inverse
kinematics adapter once per control tick, synchronously. It is
deterministic given the seed, runs faster than real time, and shares the
task scripts, environment, gating and dataset schema with the teleop
oracle, so either can produce any dataset. The collection metadata
records which oracle produced each dataset.

\paragraph{Reliability.}
With the corrections in place the direct oracle succeeds on every
attempted scenario of both tasks. The teleoperation-pipeline oracle
matches it on the single-arm task and reaches a high but imperfect
per-attempt rate on the relay (its residual failures are place-phase
misses through pipeline lag). Because collection retries a failed seed a
bounded number of times and saves only verified successes, nearly every
training seed still yields one clean demonstration, and no failed
episode can enter a dataset.
