\section{Difficulties and resolutions}
\label{sec:difficulties}

Every environment of this kind embodies dozens of small decisions made
under failure. Hiding them would make the result irreproducible: a
reader rebuilding the environment would meet the same failures with no
record of the fixes. This section catalogues each difficulty in the
order we met it --- contact, geometry, control, then tooling --- with
its symptom, diagnosis and resolution, and closes with what remains
unjustified.

\subsection{Contact grasping}
\label{sec:diff-grasp}

The first cube grasps failed roughly one attempt in three: the jaws
closed beside the cube, or clipped its edge and spun it away. The
diagnosis was not contact physics but tracking: the differential solver
carries an azimuth-dependent steady-state error of roughly
\SI{8}{\milli\metre}, so a script that positions the gripper perfectly
in its own frame places the real pinch centre off the cube. The earlier
payload, a ten-centimetre prism, had masked this error entirely.

The obvious fix --- a proportional servo centring the pinch during the
whole approach --- made matters worse, dropping success to zero. The
pipeline lags its target by several ticks, so the servo integrated stale
error into wind-up and swept the gripper past the cube. The working
form holds the approach for a convergence dwell, then corrects with a
low gain under a hard magnitude clip, and freezes the correction the
moment the jaws start closing.

The decisive improvement came from visual review of failure renders:
the fingertips hovered just above the cube's top face at the commanded
grasp depth. Commanding the pinch \emph{below} the cube's centre ---
lower than any geometrically sensible target --- lets the solver's
vertical undershoot land the pinch on the cube. With the dwell, the
clipped servo and the deepened grasp, single-arm grasping became
fully reliable.

\subsection{Place-phase dragging}
\label{sec:diff-drag}

Relay episodes then failed at the \emph{second} place: lengthening the
settle dwell --- normally the safe change --- made success \emph{worse}.
The hold servo was the culprit: after the cube touched down but before
the jaws opened, the servo kept correcting through the pipeline's lag
and dragged the placed cube off its target; more dwell meant more
dragging. The resolution freezes the hold correction once a previously
lifted cube is detected back at rest while still gripped. The general
lesson: any closed-loop correction acting through a laggy actuation
path needs an explicit end-of-life rule, not just a gain.

\subsection{Workspace envelope versus the twin table}
\label{sec:diff-envelope}

Teleoperation targets pass through an analytic workspace envelope whose
floor protects the real rig's table surface. The twin's table top sits
about \SI{34}{\milli\metre} below the plane that floor assumes, so
table-level grasp targets in simulation were silently projected upward
--- the gripper stopped short of the cube with no error anywhere. The
fix scopes an envelope-floor override to the simulated collection scene
only; the real stack's tuned floor is untouched. The implication worth
recording for the real rig: whenever the physical working surface
differs from the height the envelope was tuned at, grasps near the
surface will fail in exactly this silent way.

\subsection{The abandoned mid-air hand-over}
\label{sec:diff-handover}

The original bimanual task held the object in the air while the second
arm grasped it. Two failure modes resisted all tuning. First, a
parallel-jaw grasp leaves the object's yaw about the vertical axis
unconstrained; during the carry the object rotated freely, so the
second gripper's approach met an unpredictable cross-section. Second,
two gripper bodies converging on one small object collided with each
other. Both problems vanish when the object is staged on the table
between grasps, which is why the relay of Section~\ref{sec:tasks}
exists. The lesson generalises: with parallel jaws and no wrist
redundancy, prefer task decompositions in which every grasp starts from
a fully constrained object pose.

\subsection{Renderer cache}
\label{sec:diff-renderer}

The twin cached one offscreen renderer keyed by image width alone.
Requesting two camera resolutions silently returned buffers of the
wrong height for one of them. The cache is now keyed by the full
resolution. A one-line bug, but it cost a debugging session because the
symptom (distorted frames in one stream) appeared far from the cause.

\subsection{Training-stack pitfalls}
\label{sec:diff-tooling}

For reproducibility, the pitfalls met wiring three policy families to
locally collected datasets, none of which produced a helpful error
message at the point of failure:

\begin{itemize}
  \item \textbf{Normalisation lives outside the policy.} The framework's
    policies do not normalise inside their action-selection call;
    normalisation is a separate processor pipeline saved alongside the
    checkpoint. An evaluation harness that loads only the policy weights
    silently evaluates an unnormalised model. Ours loads the
    checkpoint's own processor pipelines and passes every observation
    and action through them.
  \item \textbf{Observation construction must match the dataloader.}
    Evaluation images must be channel-first floating point in the unit
    range, keyed exactly as the dataset features were; we reuse the
    framework's own observation-conversion helper rather than
    reimplementing it.
  \item \textbf{Output directories must not pre-exist.} The trainer
    refuses to write into an existing run directory; resumable scripts
    must reuse finished cells and clear partial ones.
  \item \textbf{Video decoding.} Dataset videos are encoded with a
    modern codec; the default decoding backend loads system libraries
    that a sudo-less machine may lack. The pure-Python backend bundles
    its own decoder and is passed explicitly everywhere.
  \item \textbf{Gated model access.} The vision-language policy's
    tokenizer lives in a licence-gated repository; an unauthenticated
    machine fails at download time deep inside training. Our scripts
    check access in preflight and either skip the cell (smoke test) or
    abort before any compute is spent (long run).
  \item \textbf{Model sizing.} Full finetuning of the
    four-billion-parameter vision-language policy does not fit a
    \SI{24}{\gibi\byte} GPU once optimiser state is counted; low-rank
    adaptation of the published base fits comfortably and is the
    configuration our long-run script uses.
\end{itemize}

\subsection{What remains unjustified}
\label{sec:diff-unjustified}

The contact constants (friction, impedance ratio, pad geometry), the
payload mass, the servo gains, clips and dwell durations, and the
oracle-gate thresholds beyond their measured bands are all tuned rather
than derived; each is flagged where introduced. They are stable across
the seeds exercised so far, and the oracle gate exists precisely to
catch a future regression in any of them before it can poison a
training run.
