\section{Experiments}
\label{sec:experiments}

The experiment asks two questions: do contemporary imitation policies
--- ACT~\citep{zhao2023aloha}, Diffusion
Policy~\citep{chi2023diffusion}, and the vision-language-action model
$\pi_{0.5}$~\citep{pi05} --- learn the two tasks from oracle
demonstrations collected through the teleoperation pipeline; and does
the simple/full mode split localise failures correctly, with simple
mode catching pipeline bugs and full mode measuring generalisation.
This section fixes the protocol; the results subsection is populated
from the long-run report as the living document grows.

\paragraph{Protocol.}
The campaign runs per mode (simple, then full), per task, per policy:
\begin{enumerate}
  \item \textbf{Oracle gate.} Both oracles are dry-run on thirty
    scenarios per task. The experiment aborts if the
    teleoperation-pipeline oracle drops below \SI{95}{\percent} on the
    single task or \SI{70}{\percent} on the relay. The relay bar sits
    below the oracle's measured per-attempt band (roughly
    \SIrange{75}{86}{\percent}) so a thirty-episode sample's noise
    cannot false-abort, yet far above the failure regime of a broken
    oracle.
  \item \textbf{Collection.} Simple mode collects on its one scenario;
    full mode collects one demonstration per training seed. Failed
    episodes are discarded and retried; only verified successes are
    saved.
  \item \textbf{Training.} ACT and Diffusion Policy train from scratch;
    $\pi_{0.5}$ is adapted from its published base with low-rank
    adapters, full finetuning being out of memory reach
    (Section~\ref{sec:diff-tooling}). All training uses the
    LeRobot framework~\citep{cadene2024lerobot}.
  \item \textbf{Checkpoint selection.} Every saved checkpoint rolls out
    on the validation seeds; the best checkpoint per cell (ties to the
    later step) advances.
  \item \textbf{Evaluation.} The selected checkpoint runs once per
    evaluation seed --- thirty held-out scenarios in full mode, thirty
    trials of the fixed scenario in simple mode --- with a video saved
    per episode. Success uses the collection criterion unchanged.
\end{enumerate}

\paragraph{Results.}
\textit{TODO: populate from the long-run report (success-rate table per
mode\,$\times$\,task\,$\times$\,policy with placement-error statistics,
per-seed success maps, validation curves, and head-to-head episode
renders) once the full campaign completes on the training machine. The
plumbing of every cell of this protocol --- collection, training,
checkpoint selection, evaluation, reporting --- has been verified end to
end at smoke scale.}
