\documentclass[10pt]{article}

\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amssymb}
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{siunitx}
\usepackage[numbers]{natbib}

% Guideline convention: flag design choices that lack a solid reason.
\newcommand{\unjustified}{\textbf{[!]}~\emph{Unjustified choice:}}

\title{A Digital-Twin Training Environment for Bimanual
Visuomotor Policies: Contact-Based Tasks, Pipeline-Authentic Oracles,
and Gated Demonstration Collection}

\author{so101\_garment project\thanks{Living document: this paper is the
canonical description of the simulation training environment in the
\texttt{so101\_garment} repository and MUST be updated, in the same
branch, whenever the simulated tasks, oracle demonstrators, collection
gating, or experiment protocol change. The repository's writing
guideline governs its style. The teleoperation system it builds on has
its own living paper in the same repository.}}

\date{Last updated: \today}

\begin{document}
\maketitle

\begin{abstract}
Imitation-learning pipelines for bimanual manipulation fail late and
expensively when their first end-to-end test uses real-robot
demonstrations: schema mismatches, normalisation errors and evaluation
bugs all surface only after days of collection. We instead prove the
whole collect--train--evaluate loop inside a digital twin of our dual
five-degree-of-freedom rig before any hardware demonstration exists. The
environment adds contact-based cube manipulation to the twin: a
single-arm pick-and-place task and a bimanual relay in which one arm
stages the object at a shared midline for the other. Demonstrations come
from a scripted oracle that drives the \emph{full production
teleoperation pipeline} --- the same input filtering, clutch
calibration, retargeting and asynchronous solver a human operator
exercises --- by inverting the retargeting map, so the collected data
distribution matches what the real rig records. Closed-loop grasp and
placement corrections, standing in for the teleoperator's eyes,
compensate the solver's steady-state tracking error. Collection is
gated: only episodes that verifiably succeed enter the dataset. A
disjoint seed protocol separates training, validation and evaluation
scenarios, and a deliberately trivial overfit-one-scenario mode
distinguishes pipeline bugs from data problems. We document every
difficulty met while building the environment --- contact tuning,
servo wind-up, envelope conflicts, tooling pitfalls --- so the result is
reproducible, and we reserve an experiments section that grows with the
living document.
\end{abstract}

\input{sections/01_intro}
\input{sections/02_environment}
\input{sections/03_tasks}
\input{sections/04_oracles}
\input{sections/05_difficulties}
\input{sections/06_experiments}
\input{sections/07_conclusion}

\bibliographystyle{plainnat}
\bibliography{references}

\end{document}
