\documentclass[10pt]{article}

\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amssymb}
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{siunitx}
\usepackage[numbers]{natbib}

% Guideline convention: flag design choices that lack a solid reason.
\newcommand{\unjustified}{\textbf{[!]}~\emph{Unjustified choice:}}

\title{Meta-Quest Teleoperation of Low-Cost 5-DoF Bimanual Arms:
Retargeting, Workspace-Envelope Handling, and a Comparison of
Inverse-Kinematics Strategies}

\author{so101\_garment project\thanks{Living document: this paper is the
canonical description of the teleoperation system in the
\texttt{so101\_garment} repository and MUST be updated, in the same
branch, whenever the teleoperation methods, orientation mapping,
workspace envelope, or benchmark results change. The repository's
writing guideline governs its style.}}

\date{Last updated: \today}

\begin{document}
\maketitle

\begin{abstract}
We present a teleoperation system that drives a pair of low-cost 5-DoF
arms from Meta Quest hand controllers, built to collect bimanual
demonstrations for garment manipulation. Three properties of such arms
make direct pose tracking a poor fit. The missing wrist joint makes full
6-D orientation tracking structurally over-constrained. The reachable
workspace is far smaller than a person's natural gestures. The
position-controlled bus servos reward smooth commands. We answer each
property in turn. A clutched retargeting scheme lets the operator work in
comfortable strokes and re-anchor at will. An orientation mapping we call
\emph{armplane} constructs only wrist orientations the arm can actually
realise. An adaptive input filter suppresses hand tremor without adding
noticeable lag. An analytic workspace envelope detects infeasible targets
and offers a choice of policies for handling them. Beneath these sits a
pluggable inverse-kinematics layer in which quadratic-programme trackers,
damped least squares, nonlinear optimisation, and an analytic split
solver are interchangeable. We evaluate every component in a digital twin
of the rig, with benchmark suites for position tracking, wrist agility,
and out-of-envelope behaviour, and we pair the benchmark with a
within-subjects user-study protocol. The appendix specifies every method
in full implementation detail.
\end{abstract}

\input{sections/01_intro}
\input{sections/02_related}
\input{sections/02b_preliminaries}
\input{sections/03_system}
\input{sections/04_method}
\input{sections/05_experiments}
\input{sections/06_conclusion}

\bibliographystyle{plainnat}
\bibliography{references}

\appendix
\input{sections/appendix_methods}

\end{document}
