\section{Related work}
\label{sec:related}

This section positions the system along four axes: how other systems
teleoperate bimanual low-cost arms, how other systems control arms with
5-DoF wrists, which inverse-kinematics families our pluggable layer
draws on, and the standard treatments of input filtering and workspace
awareness.

\paragraph{Bimanual teleoperation of low-cost arms.}
Two families dominate. Joint-space leader--follower systems pair each
follower arm with a kinematically matched leader that the operator moves
by hand: ALOHA and Mobile ALOHA drive bimanual manipulation this
way~\citep{zhao2023aloha,fu2024mobilealoha}, and GELLO generalises the
idea with low-cost scaled leader replicas for common
manipulators~\citep{wu2024gello}. Leader--follower transfers joint
angles directly, so it sidesteps inverse kinematics entirely, but it
requires dedicated leader hardware and ties the operator's posture to
the follower's kinematics. Task-space retargeting systems instead track
the operator's hands with vision or a VR headset and solve for joint
angles online: AnyTeleop retargets vision-based hand poses to arm--hand
systems~\citep{qin2023anyteleop}, and Open-TeleVision streams immersive
stereo feedback to a headset-tracked operator~\citep{cheng2024televison}.
Closest to our setting, LeVR adds a VR teleoperation path for the LeRobot
imitation-learning stack and uses it to collect headset demonstrations
that fine-tune visuomotor policies~\citep{levr}; it targets a
higher-DoF arm-and-hand system, whereas we target the under-actuated
bimanual SO-101 and the retargeting problems its 5-DoF wrist creates.
Our system takes the task-space route with Meta Quest controllers: it
needs no leader hardware, lets the operator stand and move freely, and
gains clutching and motion scaling for free~--- at the price of needing
the retargeting and IK machinery this paper describes.

\paragraph{Controlling 5-DoF wrists.}
Telegrip~\citep{telegrip} teleoperates the SO-100 lineage from WebXR
controllers with a deliberately minimal scheme: position-only IK on the
three proximal joints and a direct mapping from controller rotation
deltas to the two wrist joints, with joint limits clamped after the
solve. Its wrist feels agile precisely because the wrist path bypasses
any task-space optimisation. Our \texttt{telegrip} method
(Section~\ref{sec:telegrip}) ports this split-IK idea to our benchmark
interface, replacing its physics-engine IK with the simulator's own
kinematics and recovering the wrist angles analytically from the
absolute target orientation. Other VR pipelines for hobby-class arms
(e.g.\ the damped-least-squares strategy our \texttt{dls} baseline
reproduces) typically track the full 6-D pose and inherit the
over-constraint problem on 5-DoF hardware. Our armplane construction
(Section~\ref{sec:armplane}) takes the opposite approach: rather than
relaxing the tracker, it constructs targets inside the arm's reachable
orientation set, so any tracker can follow them.

\paragraph{Differential IK and task-space control.}
Our armplane pipeline builds on Pink~\citep{pink}, a task-based
differential-IK library over Pinocchio~\citep{pinocchio}, solving a QP
with frame-tracking, damping, and posture tasks under joint limits;
mink~\citep{mink} solves the analogous problem over MuJoCo models.
Damped least squares is the classical singularity-robust baseline
\citep{wampler1986,nakamura1986}. The benchmark also includes a per-tick
bounded nonlinear least-squares solve (\texttt{scipy\_ls}) as an
optimisation-based reference.

\paragraph{Input filtering.}
Controller pose noise is attenuated with the One-Euro
filter~\citep{oneeuro}, the de-facto standard for interactive input: a
first-order low-pass whose cutoff adapts to signal speed, trading jitter
at rest against lag during fast motion (Section~\ref{sec:prelim} defines
it). Filter parameters interact directly with wrist agility
(Section~\ref{sec:experiments}).

\paragraph{Workspace awareness.}
Commercial teleoperation stacks commonly project infeasible targets onto
a precomputed reachability approximation. We adopt the same philosophy
with an analytic annulus-plus-floor envelope derived from the arm model
(Section~\ref{sec:envelope}), and additionally compare \emph{freeze} and
\emph{velocity-shaping} alternatives, which to our knowledge have not
been compared head-to-head on low-cost 5-DoF hardware.
