\section{Retargeting method}
\label{sec:method}

This section specifies the \emph{armplane} retargeting pipeline, named
after its orientation construction. Section~\ref{sec:opframe} derives
the operator control frame, Section~\ref{sec:clutch} defines the clutch
and its position mapping, Section~\ref{sec:armplane} constructs the
orientation targets, Section~\ref{sec:engagement} states exactly what
happens to the gripper pose at the moment teleoperation engages,
Section~\ref{sec:ratchet} extends the clutch's ratcheting to unbounded
gripper roll on the finite-range wrist joint, and
Section~\ref{sec:envelope} adds the workspace envelope and its
out-of-envelope policies. Two alternative wrist strategies served by the
same pipeline close the section: the Telegrip-style split IK
(Section~\ref{sec:telegrip}) and the decoupled thumbstick wrist
interface (Section~\ref{sec:thumbstick}).

\subsection{Operator control frame (headset-anywhere)}
\label{sec:opframe}

The headset streams controller poses in a reference frame whose origin
and yaw depend on where the \emph{headset} was when the app started;
used raw, ``forward'' on the hand is only ``forward'' on the robot if
the headset is placed carefully. The pipeline therefore never controls
in the raw tracking frame: at \emph{every} clutch engagement it derives
an \emph{operator frame} from the two handle poses themselves~---
gravity gives \(z\) (OpenXR spaces are gravity-aligned), the horizontal
left\(\to\)right handle line gives \(-y\), and \(x = y \times z\) points
forward. The origin sits \SI{20}{cm} behind and \SI{20}{cm} above the
handle midpoint (roughly the operator's head, and drawn as the ``headset
centre'' marker in the simulator).
\unjustified{} the \SI{20}{cm} offsets are a visual proxy
set by intuition, not measured head geometry; they only place a marker
and do not affect control. All hand poses are re-expressed in this frame
before any control mathematics, so the headset can sit anywhere, and
moving it between grips has no effect. A unit test verifies exact
invariance of the control inputs under arbitrary yaw and translation of
the tracking frame.

\subsection{The clutch and the position mapping}
\label{sec:clutch}

\paragraph{What the clutch is.}
The \emph{clutch} is a deadman-style engage/disengage control: the robot
follows the operator's hands only while both grip buttons are held
beyond a threshold. Releasing either grip freezes the arms where they
are; gripping again re-engages tracking from the current hand position.
The name comes from the automotive clutch, which likewise couples and
decouples two systems that keep their own state while separated.
\unjustified{} the grip threshold (0.9 of full grip
travel) is a convention chosen to reject accidental grips; it has not
been tuned against user data.

\paragraph{Clutched position mapping.}
Let \(p_H(t)\) be the filtered controller position in the operator frame
and \(p_E\) the arm's end-effector (EE) position. On the rising edge of
the clutch (both grips held), the pipeline latches the hand reference
\(p_H^0 = p_H(t_0)\) and the robot anchor \(p_E^0 = p_E(t_0)\). While
clutched, the target position is the anchored frame-aligned delta
\begin{equation}
  p^\ast(t) \;=\; p_E^0 + s\,\bigl(p_H(t) - p_H^0\bigr),
  \label{eq:clutch}
\end{equation}
with translation scale \(s\) (\(=0.8\) in the current configuration; the
value is derived from the operator's height against the arm's usable
radial band, and independently matches the value found by manual tuning
--- Appendix~\ref{app:common-input} gives the derivation and flags its
anthropometric links as estimates).
Releasing the grips freezes the arms; re-gripping re-anchors, so the
operator can reposition their hands (``ratcheting'') without moving the
robot. Every re-grip is a \emph{full} recalibration~--- operator frame,
handle axes, and roll references (Section~\ref{sec:armplane}) are all
re-captured~--- so a poor initial calibration is fixed by simply
releasing and gripping again.

\paragraph{Why clutched rather than absolute.}
Against a fixed absolute mapping (Section~\ref{sec:prelim-retarget}),
clutching buys three things on this hardware. First, workspace
ratcheting: the arm's reach is far smaller than the operator's, and
clutching covers it in comfortable strokes. Second, safety and comfort:
release is an instant, always-available pause, and the operator can
stretch, re-place the headset, or walk around without robot motion.
Third, self-repair: because every engagement recalibrates everything, a
bad first calibration never persists. The costs are real but smaller.
There is no persistent hand--robot correspondence, so the mapping drifts
by design across grip cycles, and spatial reasoning must be relative
rather than absolute. Each cycle also adds a small workload, which the
user study measures directly with its clutch questionnaire item
(Section~\ref{sec:exp-userstudy}).

\subsection{Armplane orientation construction}
\label{sec:armplane}

Full 6-D orientation tracking is over-constrained on the 5-DoF SO-101:
the missing wrist-yaw joint couples EE yaw to the shoulder-pan angle, so
a naive tracker sacrifices position accuracy chasing an unreachable
orientation (Section~\ref{sec:experiments}; this failure mode motivated
the design). Instead, the armplane pipeline \emph{constructs} a target
orientation that is reachable by construction, from three decoupled
inputs:
\begin{align}
  \text{tip elevation} \quad
  \theta(t) &= \operatorname{atan2}\!\bigl(a_z(t),\,\lVert a_{xy}(t)\rVert\bigr),
  & a(t) &= R_H(t)\,\hat h,
  \label{eq:elev}\\
  \text{tip azimuth} \quad
  \phi(t) &= \operatorname{atan2}\!\bigl(p_{E,y}-b_y,\;p_{E,x}-b_x\bigr),
  \label{eq:azim}\\
  \text{roll} \quad
  \psi(t) &\;\text{from the controller ``knuckle'' axis, orthogonalised
  against the tip.}
  \label{eq:roll}
\end{align}
Here \(\hat h\) is the handle axis captured per hand at every grip (the
operator holds both handles plumb-vertical, so ``handle-down =
gripper-down'' until the next re-grip), and \(b\) is the arm's base
position: the tip azimuth \emph{follows the arm} rather than the hand,
because the arm cannot yaw its wrist independently
(Section~\ref{sec:system}). The target rotation
\(R^\ast = [\,\hat t \;\; \hat y \;\; \hat z\,]\) assembles the tip
direction \(\hat t = (\cos\theta\cos\phi,\ \cos\theta\sin\phi,\
\sin\theta)\) with a roll-consistent frame completion. All three EE
orientation axes can then be fully costed in the tracker without risking
the \(180^\circ\) tip-flip that a zero-costed yaw axis admits.

Because the azimuth follows the arm's live end-effector heading, and the
roll frame is orthogonalised against the tip, any change in that heading
also moves the commanded roll. This matters at the reach boundary: while
a target sits outside the workspace envelope
(Section~\ref{sec:envelope}) the saturated arm slides along the
boundary, and its heading would otherwise swing the roll reference with
it. The pipeline therefore \emph{freezes} the azimuth reference while a
target is out of envelope, so the commanded roll stays put instead of
drifting as the arm saturates.

\subsection{Gripper orientation at engagement}
\label{sec:engagement}

What happens to the gripper the moment teleoperation starts is worth
stating exactly, because it is where naive pipelines jerk the wrist.
On the rising edge of the clutch, the pipeline performs three captures.
First, it derives the operator frame from the two handle poses
(Section~\ref{sec:opframe}). Second, it captures each hand's
\emph{handle axis}: the operator holds both handles pointing straight
down, and the body-frame direction that is currently world-down becomes
that hand's elevation reference until the next re-grip. Third, it
captures each gripper's \emph{roll reference}: a horizontal ``knuckle''
axis taken from the gripper's \emph{current} roll, so the roll target is
anchored to wherever the gripper already is.

The commanded pose then starts from the gripper's own state, not from
the hand's. The position target is anchored at the end-effector position
at grip (Eq.~\eqref{eq:clutch} with \(p_H = p_H^0\)), so the gripper
does not translate at engagement. The orientation target is blended from
the end-effector's rotation \emph{at engagement} to the constructed
armplane target over \SI{1}{s} by geodesic interpolation (spherical
linear interpolation on the rotation group), so the wrist glides rather
than snaps even if the hand's orientation disagrees with the gripper's.
\unjustified{} the \SI{1}{s} blend duration is a comfort
value chosen by feel; it has not been tuned against user data.

This behaviour is identical for every method in the pluggable layer: the
blended, reachable target of this section is computed once, upstream of
the IK interface, and every method~--- the armplane QP tracker and all
benchmark methods alike~--- receives the same target stream. The methods
differ only in how they track it. In particular, the \texttt{telegrip}
method recovers its wrist angles analytically from this same blended
target, so its engagement behaviour matches the others even though its
steady-state tracking differs (Appendix~\ref{app:methods}).

\subsection{Roll ratcheting and the wrist-roll limit}
\label{sec:ratchet}

Because the roll reference is re-captured from the gripper's current
roll at every engagement (Section~\ref{sec:engagement}), repeated
twist--release--untwist--re-grip cycles accumulate gripper roll: the
clutch ratchets orientation exactly as it ratchets position. What stops
the accumulation is the joint itself. The wrist-roll joint spans only
about \(320^\circ\), so a few cycles in one direction park it at a
limit, where further hand twist does nothing and the operator perceives
the roll as stuck~--- a real obstacle when aligning the gripper to grasp
or hand over an elongated object.

The pipeline removes this ceiling with a \emph{jaw-equivalence rewrap}.
A parallel-jaw gripper is functionally identical under a half-turn of
roll: the jaws swap, and every grasp it could make it can still make.
When an engagement finds the solved wrist-roll within a guard band of a
joint limit, the engagement re-anchors the roll reference to the
jaw-equivalent orientation half a turn away, restoring headroom at no
functional cost; the activation blend (Section~\ref{sec:engagement})
glides the wrist there, so nothing snaps. The rewrap never fires while
the trigger is held~--- an object in the jaws is never spun uninvited
--- and it only ever happens at an engagement edge, never mid-hold.
While a hold sits inside the guard band, the operator instead receives a
throttled reminder to release, untwist, and re-grip. A per-hand joystick
click queues a \emph{roll reset}: the next engagement anchors the roll
at the joint's neutral and the blend glides the gripper back.
\unjustified{} the guard band (\(8^\circ\)) is a convention balancing
early intervention against usable range; it has not been tuned against
user data. The jaw-equivalence assumption also fails if the roll
attitude of a wrist-mounted camera matters to the task; no such
constraint exists on this rig today.

\subsection{Workspace envelope and out-of-envelope policies}
\label{sec:envelope}

\paragraph{Envelope geometry.}
The reachable position set of one arm is approximated by an annulus
around the shoulder-lift pivot plus a table-clearance floor:
\begin{equation}
  \mathcal{W} \;=\; \bigl\{\,p \;:\; r_{\min}\le \lVert p - c(p)\rVert \le
  r_{\max}, \;\; p_z \ge z_{\mathrm{floor}} \,\bigr\},
  \label{eq:envelope}
\end{equation}
where the pivot \(c(p)\) is \emph{not} on the pan axis: it orbits it at
\SI{35.5}{mm}, and is recomputed from the target's azimuth about the pan
axis (the pan angle tracks the target azimuth in steady state; the
transient error is absorbed by a \SI{10}{mm} safety margin). Because the
lift joint rotates the entire distal chain rigidly about the pivot, the
pivot-to-EE distance is invariant to the lift angle; the radii are
obtained by a forward-kinematics grid sweep over elbow and wrist-flex
within joint limits, giving \(r_{\min}=\SI{0.084}{m}\),
\(r_{\max}=\SI{0.411}{m}\) (a unit test re-derives them so model edits
cannot silently stale the constants). Point membership, signed margin,
and nearest-point projection are a handful of vector operations
(\(\approx\)\SI{36}{\micro s}), computed every tick at \SI{100}{Hz} for
both arms. The floor \(z_{\mathrm{floor}}\) sits a few millimetres
\emph{below} the table plane rather than above it: table-contact grasps
need the gripper to actually reach the surface, and stopping the target
just short of it would leave a persistent gap. The small allowed
penetration is bounded by the same per-tick clamp, so the arm never
drives hard into the table.

\paragraph{Policies.}
Four selectable policies decide what reaches the IK when the operator's
target \(p^\ast\) leaves \(\mathcal{W}\):
\begin{description}
  \item[warn] Pass \(p^\ast\) through unchanged; log and warn (legacy
    behaviour plus telemetry). The QP saturates against joint limits.
  \item[project] Emit \(\Pi_{\mathcal{W}}(p^\ast)\), the nearest envelope
    point. The arm slides along the boundary, staying as close to the
    hand as physically possible; re-entry is seamless since the emitted
    target is always consistent with the arm's position.
  \item[freeze] Hold the last feasible target while outside (``leash'');
    release on re-entry. State resets when the clutch releases.
  \item[slow] Velocity shaping: the per-tick step \(\Delta p\) is split
    into the outward-normal component (w.r.t.\ the nearest boundary) and
    the tangential component; the outward part is scaled by
    \(\sigma\bigl(m/m_{\mathrm{soft}}\bigr)\) where \(m\) is the current
    margin, \(m_{\mathrm{soft}}=\SI{40}{mm}\), and \(\sigma\) is the
    cubic smoothstep~--- full speed far inside, zero at the boundary,
    tangential sliding unaffected. If a fast hand still escapes, the
    result degrades to projection.
\end{description}
The policy runs in the IK thread \emph{after} the height lock (so a
height locked below the floor is still clamped) and applies to every
input source (headset, mock device, interactive gizmos). The mode is
selectable at launch on both the real-robot and simulation tools.

\paragraph{Operator feedback.}
Whichever policy is active, the operator should also \emph{feel} that
their hand has left the envelope, because the visual cue (the arm
stopping or sliding) is easy to miss mid-task. A feedback layer
therefore observes the per-arm envelope status each tick and emits
debounced cues: one cue immediately when a target crosses outside, a
repeat at a fixed period while it stays outside, and a single stop cue
on re-entry. Cue intensity grows smoothly with penetration depth,
saturating one soft-margin outside the boundary, so a grazing excursion
feels different from a deep one. The current implementation signals
audibly through the control computer: a terminal bell plus a margin
readout, and~--- since a terminal is easy to mute or look away from~--- a
real speaker tone played on the machine's default audio device. The two
arms get different pitches (a lower tone for the left arm, a higher one
for the right), so the operator hears \emph{which} arm has left the
envelope without reading anything. Controller vibration is still the
intended endpoint, and the transport for it is specified~--- a
one-way-to-two-way extension of the existing headset link~--- but awaits
a rebuild of the headset application.
\unjustified{} the repeat period (\SI{0.25}{s}) is a convention chosen
to be noticeable without being irritating; it has not been tested with
users.

\subsection{Telegrip-style split IK}
\label{sec:telegrip}

The \texttt{telegrip} method ports Telegrip's split control
scheme~\citep{telegrip}: only the three proximal joints solve an IK
problem, and they solve it for \emph{position only}; the wrist joints
are set directly, bypassing every task-space cost trade-off. Per tick
and per arm:
\begin{enumerate}
  \item \textbf{Analytic wrist.} Because the lift/elbow/flex axes are
  parallel (Section~\ref{sec:system}), the EE tip elevation is affine in
  those joints,
  \(\theta = e_0 + s_2 q_2 + s_3 q_3 + s_4 q_4\) with slopes
  \(s_i\in\{\pm1\}\); given the currently commanded \(q_2,q_3\), the
  wrist flex realising the target elevation \(\theta^\ast\) is
  \(q_4 = s_4(\theta^\ast - e_0 - s_2 q_2 - s_3 q_3)\). Similarly the
  roll angle of the EE's local \(z\) about the tip axis (measured
  against a horizontal reference) is affine in the roll joint,
  \(\psi = r_0 + s_5 q_5\). The coefficients are \emph{calibrated
  numerically} by finite differences at construction rather than derived
  from the arm model (whose wrist-roll frame carries a non-trivial
  rotation that makes sign bookkeeping treacherous); a unit test
  validates the affine model to \(<1^\circ\) across the joint range.
  When the tip is near-vertical the roll reference degenerates and the
  previous roll is held.
  \item \textbf{Position DLS.} With the wrist frozen, a damped
  least-squares step on the \(3\times3\) position Jacobian of the
  proximal joints,
  \(\Delta q = J^\top (J J^\top + \lambda^2 I)^{-1}(p^\ast - p)\)
  with \(\lambda=0.05\) and per-tick gain \(0.6\).
  \item \textbf{Clamping.} All five joints are velocity-clamped
  (\SI{3}{rad/s}) and position-clamped to their limits.
\end{enumerate}
The method registers in the same pluggable layer as the QP baselines and
is selectable everywhere. Appendix~\ref{app:telegrip} details the full
port and its deliberate departures from the upstream implementation.

\subsection{Decoupled thumbstick wrist interface}
\label{sec:thumbstick}

Both wrist strategies so far derive the gripper attitude from the
operator's hand: the armplane mapping from absolute hand orientation, the
Telegrip split IK from the same constructed target. On the real rig, fine
control of wrist roll and flex through absolute hand orientation proved
hard (operator report): small, deliberate twists of the wrist are exactly
where the missing yaw joint couples attitude into position, and where
hand tremor is most visible. This interface takes the opposite route. It
keeps the relaxed position solver (Section~\ref{sec:exp-tracking}) for
the hand-driven position target, and moves wrist roll and flex onto the
headset thumbsticks as \emph{joint-space rate trims}. Because a trim
commands a joint rate directly, it never competes with a task-space cost
and never excites the wrist-axis coupling.

\paragraph{The trims.}
Each controller's thumbstick drives that arm's wrist: the horizontal axis
trims wrist roll and the vertical axis trims wrist flex. Each axis is
shaped by a deadzone and an exponential curve~--- zero inside the
deadzone, then rising smoothly from zero at its edge to a full rate at
full deflection~--- and the shaped value is integrated in joint space at
that rate. The trims are themselves \emph{clutched}, in the same
deadman sense as the grip clutch (Section~\ref{sec:clutch}) but per
stick: while a stick sits beyond its deadzone, that arm's other joints
are held exactly where they were when the stick engaged and the handle is
ignored, so only the wrist moves; the moment the stick returns to centre
the wrist stays where it was left. A forward-kinematics guard protects
the table: before a flex increment is applied, the pipeline runs the
candidate configuration through forward kinematics and rejects the
increment if it would drop the tip below the floor (roll cannot lower the
tip, so it is never blocked). On release the hand-to-robot correspondence
re-anchors against the trimmed pose~--- the position delta is zeroed and
the attitude reference is reset to what the sticks left~--- so handle
control resumes from the new wrist attitude without a jump.

\paragraph{Attitude between trims.}
Three behaviours govern the gripper attitude while no stick is deflected,
selectable by the operator. \emph{Hold} (the default) freezes the
attitude captured at grip, so the sticks are the only thing that ever
changes it~--- the steadiest option, and the reason the position solver
can stay fully relaxed. \emph{Incremental} keeps the wrist-motion mapping
of Section~\ref{sec:prelim-retarget} and lets the trims add on top.
\emph{Soft} follows the hand attitude weakly, so a trim decays back
toward the hand over the following seconds rather than persisting.

\paragraph{Trade-offs against the armplane mapping.}
The interface buys two things. Clutched precision: a stick trim is a slow,
bounded joint rate, so the operator can dial in a wrist angle far more
finely than by twisting a tracked hand. Start-anywhere: in hold mode the
attitude is anchored at the gripper's own state at grip, so teleoperation
engages from any hand pose without a wrist snap. The cost is the direct
hand-attitude correspondence that armplane preserves: outside soft mode
the gripper no longer mirrors the hand's orientation, so the operator
must track the wrist attitude through the sticks and the rehearsal triads
(Section~\ref{sec:system}) rather than proprioception. The full
algorithm, including the shaping law and the re-anchoring, is in
Appendix~\ref{app:mymethod}.
\unjustified{} the deadzone (0.15 of full throw), the roll and flex
rates (\(60^\circ\)/s and \(45^\circ\)/s at full deflection), and the
expo exponent (2.0) are initial engineering guesses, not yet tuned
against the real controllers; and the two trim direction signs are
unverified on the real robot, so a reversed axis is a one-line
configuration change rather than a code fix.
