\section{Implementation details of all teleoperation methods}
\label{app:methods}

This appendix specifies every teleoperation method in the repository in
full implementation detail: the algorithm and update law, every
parameter with its value and the reason for it (or an explicit
unjustified-choice flag), the initialisation and reset behaviour, and
the known failure modes. Together with Section~\ref{sec:method}, it is
intended to be sufficient to reimplement the whole method suite without
reading our code.

\paragraph{Common interface and rate limiter.}
All methods implement one narrow interface: reset to a given joint
configuration, then per tick accept absolute EE pose targets (robot base
frame, both arms) and return a 10-DoF joint command in a fixed joint
order. An adapter exposes any of them behind the real-time solver
interface and applies a joint-space rate limiter
\(\lvert\Delta q\rvert \le \dot q_{\max}\,\Delta t\) after every solve
(\(\dot q_{\max}=3\)~rad/s in simulation, 2~rad/s on hardware). The
limiter exists because the benchmark showed \texttt{scipy\_ls} and
\texttt{mink} can jump between IK branches near the workspace edge; the
limiter bounds the jump rate. \unjustified{} the two
\(\dot q_{\max}\) values are safety conventions (hardware stricter than
simulation), not derived from servo ratings. Grippers are excluded from
IK everywhere (locked in the reduced models).

\subsection{\texttt{armplane} / \texttt{pink\_full} — QP tracker, full orientation cost}
\label{app:pink-full}

\emph{Algorithm.} Differential IK with Pink over a reduced Pinocchio
model (grippers locked). Tasks: a 6-D frame task per EE (position cost
1.0, orientation cost 0.75, task gain 0.4, Levenberg--Marquardt damping
0), a damping task (cost 0.25), and a per-joint-weighted posture task
(zero cost by default). The QP solves for joint velocities under
configuration and velocity limits (solver \texttt{quadprog}, Tikhonov
damping \(10^{-12}\)), then integrates with \(\Delta t = 10\)~ms.

\emph{Parameters.} Position cost 1.0 is the reference scale. Orientation
cost 0.75 keeps all three orientation axes firmly tracked, which is safe
\emph{only} because the armplane construction feeds this tracker
reachable orientations (Section~\ref{sec:armplane}); on raw 6-D targets
the same weight trades position for orientation error
(Section~\ref{sec:exp-tracking}). Task gain 0.4 is a low-pass on target
tracking: lower is smoother but laggier.
\unjustified{} the 0.75 orientation cost, 0.4 gain, 0.25
damping cost, and the zero posture cost were set by iterative manual
tuning on the rig; they work well but carry no derivation.

\emph{Initialisation/reset.} Anchors to the measured joints; at clutch
engagement it receives the blended targets of
Section~\ref{sec:engagement}, so no solver-side smoothing is needed.

\emph{Failure modes.} As \texttt{pink\_full} in the benchmark (raw 6-D
targets), it exhibits the over-constraint penalty: position error grows
whenever the commanded orientation is unreachable. As the
\texttt{armplane} tracker (constructed targets) this mode does not
arise.

\subsection{\texttt{pink\_relaxed} — QP with relaxed orientation}
\label{app:pink-relaxed}

\emph{Algorithm.} Identical QP with the orientation cost lowered to
0.05: orientation becomes a soft preference and position dominates.

\emph{Parameters.} 0.05 was chosen as roughly the largest weight at
which position error stays within a few millimetres of the
position-only optimum on the tracking suite.
\unjustified{} the specific value 0.05 (rather than, say,
0.02 or 0.1) was picked from a coarse manual sweep, not an optimisation.

\emph{Initialisation/reset and failure modes.} As
Appendix~\ref{app:pink-full}. Best all-round tracker of the original
benchmark and the recommended setting when paired with the armplane
mapping. Its orientation error is bounded but non-zero by construction;
tasks needing exact wrist attitude favour \texttt{telegrip}.

\subsection{\texttt{dls} — damped least squares}
\label{app:dls}

\emph{Algorithm.} Per arm and per tick, the 6-D task error
\(e = [\,p^\ast - p;\ w_o \log(R^\ast R^\top)\,]\) (orientation weight
\(w_o=0.5\)) and the site Jacobian \(J\) (weighted likewise) give
\(\Delta q = J^\top (J J^\top + \lambda^2 I_6)^{-1} e\) with
\(\lambda = 0.05\), applied with per-tick gain 0.4, velocity clamp
3~rad/s, and joint-range clamp. Runs directly on the MuJoCo model, so it
also cross-checks the Pinocchio-based methods against the simulator's
kinematics.

\emph{Parameters.} \(\lambda=0.05\) is the classical
singularity-robustness trade-off (larger damps harder near
singularities but slows convergence); gain 0.4 matches the QP task gain
for comparability. \unjustified{} \(w_o=0.5\) follows the
community implementation this baseline reproduces, without independent
tuning.

\emph{Initialisation/reset.} Stateless beyond the current configuration;
resets to the given joints.

\emph{Failure modes.} Inherits the 6-D over-constraint problem on raw
targets; near-singular postures converge slowly rather than jumping.

\subsection{\texttt{mink} — MuJoCo-native QP}
\label{app:mink}

\emph{Algorithm.} The mink library's differential IK over the MuJoCo
model: per-EE 6-D frame tasks with the same cost ratio as
Appendix~\ref{app:pink-full} (position 1.0, orientation 0.75, gain 0.4),
a small posture cost (\(10^{-3}\)), solved as a QP with joint limits and
a velocity cap matching \texttt{dls} (3~rad/s).

\emph{Parameters.} Deliberately mirrors the Pink tracker so that
differences isolate the modelling backend (MuJoCo vs Pinocchio) rather
than the weights. \unjustified{} the \(10^{-3}\) posture
cost is a stabiliser copied from library examples.

\emph{Initialisation/reset.} Anchors to the given joints.

\emph{Failure modes.} Behaviourally close to \texttt{pink\_full},
including the over-constraint penalty; can jump branches near the
workspace edge, so the adapter's rate limiter applies.

\subsection{\texttt{scipy\_ls} — bounded nonlinear least squares}
\label{app:scipy-ls}

\emph{Algorithm.} Per tick, a bounded nonlinear least-squares solve
(Trust Region Reflective) over the arm joints, minimising the stacked
position and weighted orientation residuals (orientation weight 0.1),
warm-started from the previous solution, capped at 25 residual
evaluations per tick.

\emph{Parameters.} The evaluation cap bounds the per-tick latency; the
low orientation weight prioritises position like
\texttt{pink\_relaxed}. \unjustified{} both the 0.1 weight
and the 25-evaluation cap were set by trial to keep solve times
interactive.

\emph{Initialisation/reset.} Warm start from the previous solution; on
reset the guess is the given configuration.

\emph{Failure modes.} The most accurate position tracker in the sweep
but the most expensive per tick, and consecutive solves can land on
different IK branches near the workspace edge~--- the adapter's rate
limiter is mandatory, and residual jerk remains the highest of the QP
family.

\subsection{\texttt{telegrip} — split IK, direct wrist}
\label{app:telegrip}

\emph{Algorithm.} Port of Telegrip's control scheme
(Section~\ref{sec:telegrip}). Wrist first: \(q_4\) from the affine
elevation model \(\theta = e_0 + s_2 q_2 + s_3 q_3 + s_4 q_4\) evaluated
with the currently commanded \(q_2, q_3\); \(q_5\) from the affine roll
model \(\psi = r_0 + s_5 q_5\) (previous roll held when the tip is
within \(5.7^\circ\) of vertical). Then position: DLS on the
\(3\times3\) position Jacobian of \(q_1..q_3\) with \(\lambda=0.05\),
per-tick gain 0.6. All joints velocity-clamped (3~rad/s) and
range-clamped.

\emph{Parameters.} The affine coefficients are calibrated by finite
differences on the simulator's kinematics at construction; unit tests
validate the slopes to be exactly \(\pm1\) and the affine models to
\(<1^\circ\) across the joint range. Gain 0.6 (higher than the 6-D
methods' 0.4) is affordable because the 3-DoF position subproblem is
well-conditioned. \unjustified{} the \(5.7^\circ\)
(0.1~rad) vertical-degeneracy threshold and the 0.6 gain were set by
trial.

\emph{Initialisation/reset.} Calibration runs once at construction;
reset re-anchors to the given joints. At clutch engagement it receives
the same blended targets as every other method
(Section~\ref{sec:engagement}).

\emph{Failure modes.} The direct wrist mapping happily commands steps up
to the velocity clamp, so command jerk is an order of magnitude above
the QP trackers (Section~\ref{sec:exp-wrist}); near-vertical tips freeze
the roll axis by design; and because the wrist ignores position
consequences, fast wrist motion perturbs the tip position for the
proximal DLS to correct one tick later.

\emph{Departures from upstream Telegrip.} The physics-engine IK is
replaced by the benchmark scene's kinematics; the wrist angles are
recovered from the absolute target rotation rather than integrated from
controller deltas (equivalent under the clutch mapping, but stateless);
and the shoulder pan participates in the position DLS rather than being
wrapped separately~--- the \(\pm110^\circ\) pan range makes the
differential step's implicit wrapping sufficient.

\subsection{Common input processing (all methods)}
\label{app:common-input}

Every method receives targets through the same input processing:
One-Euro filtering of controller transforms
(\(f_{c,\min}=0.8\)~Hz, \(\beta=5.0\), \(f_{c,d}=0.9\)~Hz; position and
quaternion filtered separately with a sign-consistency guard); grip
threshold 0.9 on both hands for the clutch; per-session handle-axis
calibration at first grip; the engagement behaviour of
Section~\ref{sec:engagement} (position anchored at the EE, orientation
blended over 1~s); optional height lock (freezes target \(z\) per arm)
and mirror mode (swaps and laterally reflects the hands); and the
workspace-envelope policy (Section~\ref{sec:envelope}) applied after the
height lock, immediately before targets reach the IK.
\unjustified{} the One-Euro values (0.8, 5.0, 0.9) were
tuned by feel on the rig; a principled sweep against the wrist-agility
suite remains future work.
